Another Look at Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-04-15 reports/paper-rda.Rnw
Major update to simulation results summary. Closes #3, #4, #6, #7, and #8.

\title{A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification}

@@

<<setup, include=FALSE, cache=FALSE, echo=FALSE, warning=FALSE>>=
opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE, warning = FALSE,
               cache = FALSE, echo = FALSE)

setwd("..")
library(ProjectTemplate)
load.project()

# Clean up results for display
results <- na.omit(results)
results$d <- as.factor(results$d)
results$data_set <- as.character(results$data_set)

@@

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer, but the analysis of microarray data often presents difficult challenges
because the number of gene expression levels is typically much greater than the
sample size. Consequently, several sparse and regularized classifiers have been
proposed, often with strict covariance assumptions or eigenvalue adjustments to
ensure that the classifiers are well-posed. Here, we consider an improved
version of the regularized discriminant analysis (\emph{RDA}) classifier
popularized by \cite{Friedman:1989tm}, which requires computationally expensive
matrix operations and involves a time-consuming model-selection procedure,
making the \emph{RDA} classifier impractical for high-dimensional data. We
propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that
incorporates a more general covariance-matrix regularization approach that
includes several proposed sparse and regularization methods as special
cases. Furthermore, we overcome the computational shortcomings of the \emph{RDA}
classifier by integrating a novel dimension-reduction technique that enables a
computationally efficient model-selection procedure. Moreover, we demonstrate
with several high-dimensional microarray data sets that our proposed classifier
is competitive and often superior to several recently proposed sparse and
regularized classifiers in terms of classification accuracy. Finally, we provide
an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R
package, available on CRAN.

@@

Here, we propose a generalization of the popular \emph{RDA} classifier of
\cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix
estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally
incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA}
classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA}
classifier can attain improved classification accuracy, compared to the
\emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original
form, the \emph{RDA} classifier is impractical when $p \gg N$ because the
\emph{RDA} decision rule requires the calculation of the inverses and
determinants of the $p \times p$ covariance-matrix estimators for each
class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are
selected via cross-validation, which adds a substantial computational burden to
the model-selection process. The \emph{RDA} classifier is ill-advised on modern
data sets because the model selection requires $O(Kp^3)$ calculations for each
tuning-parameter pair.

@@

We show that the \emph{GRDA} classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
\emph{RDA} classifier and show that the sum related to the null space of the
pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of the \emph{RDA} classifier and provides a substantial increase in the
computational efficiency of the \emph{RDA} classifier and its model
selection. In fact, we show that the inverse and determinants can be reduced to
calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the
$k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly
advantageous because $n_k$ is often quite small, say, 15--20, in microarray data
sets. Thus, through a dimension-reduction technique we demonstrate that the
calculation of the \emph{RDA} decision rule involving inverses and determinants
of extremely large covariance matrices is equivalent to a decision rule
involving inverses and determinants for much lower-dimensional matrices.

@@

Here, we define our generalization of the \emph{RDA} classifier by first
simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where
$\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the convex
combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 <
\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm\Sigma}$. For brevity, for $\lambda = 1$, we define $\bm X =
[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.

@@

\begin{lemma}\label{lemma:RDA-constant-term}
Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
  1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$, $1 \le k,
k' \le K$.
\end{lemma}
\begin{proof}
Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le
K$. Then, $\xbar_k - \xbar_{k'} = \sum_{i=1}^N \beta_i \bm x_i$, where $\beta_i
= (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i = k') n_k \}$.  Then, $\bm U_2'
(\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$ is equivalent to $\bm
U_2'(\xbar_k - \xbar_{k'}) = \sum_{i=1}^N \beta_i \bm U_2' \bm x_i$. Recall that
$\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that $\bm U_2'
\in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm
  \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$,
which implies that $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
\end{proof}

@@

We used version 3.0.0 of the open-source statistical software {\tt R} for the
classification study. For each classifier's {\tt R} implementation, we used the
default settings with the exception that we explicitly set the prior
probabilities as equal. In our {\tt R} implementation of the \emph{GRDA}
classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1,
\ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant
candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
cross-validation \citep{Hastie:2008dt}.

@@

For the microarray data sets, we computed error-rate estimates for the competing
classifiers after reducing the dimension of the data sets with feature selection
to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
\cite{Xu:2009fl}. To compare misclassification rates between different values of
$d$, we applied each value of $d$ considered to the same random allocation of
training and test data. Hence, for a given random partition, the set of features
selected for a small value of $d$ is a proper subset of the features selected
for a larger value of $d$. We report the error-rate estimates along with
approximate standard errors in parentheses in Tables \ref{tab:results-d100} --
\ref{tab:results-d1000} along with the reduced dimension $d$.

The \emph{GRDA} classifier was superior in classification performance for the
majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, the
\emph{GRDA} classifier outperformed its competitors for all values of $d$
examined. For the Nakayama data set, the \emph{GRDA} classifier was outperformed
by the Dudoit, Pang, and Tong classifiers for small values of $d$ but attained
classification rates superior to its competitors for larger values of
$d$. Furthermore, Figure \textbf{TODO} emphasizes two strengths of the
\emph{GRDA} classifier: reliability of the classifier for small $d$ and
stability of the classifier for large $d$. In contrast to the other classifiers,
the \emph{GRDA} classifier exhibited less variability its misclassification
estimates for small values of $d$. Also, the \emph{GRDA} classifier yielded
superior classification accuracy as $d$ increased and was hardly effected by the
inclusion of more variables in the data sets. We attribute the apparent
stability of the error-rate estimates of the \emph{GRDA} classifier for large
values of $d$ to its novel dimension-reduction technique.

@@

The popular classifier from Guo is most competitive over large values of $d$ for
the Shipp and Singh data sets. However, for the Burczynski and Nakayama data
sets, the implementation of Guo's method in the {\tt rda} package often
classified every test observation into a single class, thus yielding larger
error rates than its competitors. Furthermore, the Guo classifier exhibited
unreliable classification performance for smaller values of $d$. The Guo
classifier required a larger subset of the features to perform competitively in
our study.

As seen in Figure ?, the \emph{GRDA} classifier exhibits an interesting bathtub
shape in selection of the lambda tuning parameter through cross-validation. (Not
sure what else to write here.)

\[ \left[\text{Table 2 goes about here }\right]\]

\[ \left[\text{Table 3 goes about here }\right]\]

\[ \left[\text{Table 4 goes about here }\right]\]

\[ \left[\text{Table 5 goes about here }\right]\]

@@

The analysis of high-dimensional microarray data sets is important to modern
biomedical studies and presents challenges with supervised classification
because the number of gene expressions greatly outnumbers the number of
collected responses. Consequently, regularization techniques are often employed
to stabilize popular classifiers for improved classification performance. We
have provided a generalization of the well-known \emph{RDA} classifier from
\cite{Friedman:1989tm}, aimed for small-sample, high-dimensional
classification. The covariance-matrix regularization employed in the \emph{GRDA}
classifier includes a large family of shrinkage methods. Our formulation of the
\emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample
covariance matrix indicates that eigenvalue-adjustment methods applied to the
\emph{LDA} classifier explains the similiarities in classification accuracy of
several regularization methods applied to the \emph{LDA} classifier reported by
\cite{Ramey:2013ji} in a high-dimensional classification study. We have provided
an invariance property that extends results of \cite{Ji:2008wp} and
\cite{Ye:2006jm}, who have provided a similar invariance property for the
\emph{LDA} decision rule.

We have provided an implementation of the \emph{GRDA} classifier in the {\tt
  regdiscrim} R package, available on CRAN. Our implementation facilitates a
rapid classification of high-dimensional data by employing the novel results
provided in this paper. Moreover, our 10-fold cross-validation procedure
typically requires a couple of seconds for data sets having large $p$. When
preliminary variable selection is applied to the data as we performed in our
classification study, the time to perform model selection is nearly
instantaneous, confirming the computational complexity reported regarding the
operations employed in the \emph{GRDA} classifier.

Although several variants of the \emph{LDA} classifier have been proposed for
high-dimensional, classification, the \emph{QDA} classifier has largely been
ignored, in part due to the large number of parameters that must be
estimated. Our proposed classifier ...

Our formulation of covariance-matrix regularization employed in the \emph{GRDA}
classifier includes a large family of shrinkage methods, such as the methods
from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other
shrinkage methods studied by \cite{Ramey:2013ji} and \cite*{Xu:2009fl}.

@@

<<microarray_results, results="asis">>=
results_summary <- ddply(results, .(data_set, d, Classifier), summarize,
                         Avg_Error = mean(Error),
                         SE_Error = sqrt(mean(Error) * (1 - mean(Error)) / length(Error)))

# Tables containing the classification results for each value of 'q'
results_summary$data_set <- Hmisc:::capitalize(results_summary$data_set)
results_summary$d <- as.factor(results_summary$d)
results_summary$Avg_Error <- round_digits(results_summary$Avg_Error, 3)
results_summary$SE_Error <- round_digits(results_summary$SE_Error, 3)
results_summary$Error <- with(results_summary, paste0(Avg_Error, " (", SE_Error, ")"))
results_summary <- subset(results_summary, select = -c(Avg_Error, SE_Error))
colnames(results_summary) <- c("Dataset", "d", "Classifier", "Error")

results_tables <- tapply(seq_along(results_summary$d), results_summary$d, function(i) {
  dcast(results_summary[i, ], Dataset ~ Classifier, value.var = "Error")
})

>>>>

\title{Another Look at Regularized Discriminant Analysis for High-Dimensional Classification}

@@

<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
# Round digits throughout document including \Sexpr statements.
# See Yihui's response here:
# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
options(digits = 3)

opts_chunk$set(fig.align = 'default', dev = 'png', message = FALSE,
               warning = FALSE, cache = TRUE, echo = FALSE, autodep = TRUE,
               fig.width = 10, fig.height = 10, out.width = "\\linewidth")
@

<<setup>>=
setwd("..")
library(ProjectTemplate)
load.project()

results <- lapply(results, function(res) {
  res <- lapply(res, function(res_d) {
    errors <- do.call(cbind, res_d$errors)
    res_d[["errors"]] <- NULL
    cbind(do.call(cbind, res_d), errors)
  })
  do.call(rbind, res)
})
results <- data.frame(do.call(rbind, results), stringsAsFactors = FALSE)

results$lambda <- as.numeric(results$lambda)
results$gamma <- as.numeric(results$gamma)
results$Dudoit <- as.numeric(results$Dudoit)
results$GRDA <- as.numeric(results$GRDA)
results$Guo <- as.numeric(results$Guo)
results$Pang <- as.numeric(results$Pang)
results$Tong <- as.numeric(results$Tong)
results$Witten <- as.numeric(results$Witten)

results$d <- factor(results$d, labels = sort(unique(as.numeric(results$d))))

# Summary of the classification error rates for each data set and number of
# variables selected, d.
error_rates <- subset(results, select = -c(lambda, gamma))
m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
                     Avg_Error = mean(Error), Num_Reps = length(Error))
error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
error_rates <- subset(error_rates, select = -Num_Reps)

# Candidate values of tuning parameters.
seq_lambda <- seq(0, 1, by = 0.05)
seq_gamma <- 10^seq.int(-1, 5)

# Summary of the optimal tuning-parameter estimates for each data set and number
# of variables selected, d.
grda_optimal <- subset(results, select = c(data_set, d, lambda, gamma))
grda_optimal$lambda <- factor(grda_optimal$lambda, levels = seq_lambda)
grda_optimal$gamma <- factor(grda_optimal$gamma, levels = seq_gamma)

grda_optimal <- ddply(grda_optimal, .(data_set, d), function(x) {
  x <- subset(x, select = c(lambda, gamma))
  melt(table(x))
})

@@

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer, but the analysis of microarray data often presents difficult challenges
because the number of gene expression levels is typically much greater than the
sample size. Consequently, several sparse and regularized classifiers have been
proposed, often with strict covariance assumptions or eigenvalue adjustments to
ensure that the classifiers are well-posed. Here, we consider an improved
version of the regularized discriminant analysis (\emph{RDA}) classifier
popularized by \cite{Friedman:1989tm}, which requires computationally expensive
matrix operations and involves a time-consuming model-selection procedure. We
propose a novel modification of the \emph{RDA} (\emph{GRDA}) classifier that
provides a straightforward interpration and employs a novel dimension-reduction
technique to enable a computationally efficient model-selection procedure, thus
overcoming the computational shortcomings of the \emph{RDA}
classifier. Moreover, we demonstrate with several high-dimensional microarray
data sets that our proposed classifier is competitive and often superior to
several recently proposed sparse and regularized classifiers in terms of
classification accuracy. Finally, we provide an implementation of the
\emph{GRDA} classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

Here, we propose a generalization of the popular \emph{RDA} classifier of
\cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix
estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally
incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA}
classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA}
classifier can attain improved classification accuracy, compared to the
\emph{LDA} and \emph{QDA} classifiers when $p < N$. However, in its original
form, the \emph{RDA} classifier is impractical when $p \gg N$ because the
\emph{RDA} decision rule requires the calculation of the inverses and
determinants of the $p \times p$ covariance-matrix estimators for each
class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are
selected via cross-validation, which adds a substantial computational burden to
the model-selection process. The \emph{RDA} classifier is ill-advised on modern
data sets because the model selection requires $O(Kp^3)$ calculations for each
tuning-parameter pair.

@@

We show that the \emph{GRDA} classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
\emph{RDA} classifier and show that the sum related to the null space of the
pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of the \emph{RDA} classifier and provides a substantial increase in the
computational efficiency of the \emph{RDA} classifier and its model
selection. In fact, we show that the inverse and determinants can be reduced to
calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the
$k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly
advantageous because $n_k$ is often quite small, say, 15--20, in microarray data
sets.

@@

Here, we define our generalization of the \emph{RDA} classifier by first
simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where
$\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the convex
combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 <
\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm\Sigma}$. In Figure \ref{fig:grda-contours} we demonstrate the
effect of $\lambda$ on the contours of five multivariate normal populations with
unequal covariance matrices. For brevity, when $\lambda = 1$, we define $\bm X =
[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$.

\[ \left[\text{Figure \ref{fig:grda-contours} goes about here }\right]\]

@@

\begin{lemma}\label{lemma:RDA-constant-term}
Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
  1}$, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$, $1 \le k,
k' \le K$.
\end{lemma}
\begin{proof}
Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
$\bm U_2' \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{\bm
  \Sigma})$, $i = 1, \ldots, N$, $\bm U_2' \bm x_i = \bm 0_{(p-q) \times 1}$.
Hence, $\bm 0_{(p-q) \times 1} = \sum_{i=1}^N \beta_i \bm U_2'\bm x_i = \bm U_2'(\xbar_k -
\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
k') n_k \}$. Therefore, $\bm U_2' (\bm x - \xbar_k) = \bm U_2' (\bm x - \xbar_{k'})$.
\end{proof}

@@

We used version 3.0.0 of the open-source statistical software {\tt R} for the
classification study. For each classifier's {\tt R} implementation, we used the
default settings with the exception that we explicitly set the prior
probabilities as equal. In our {\tt R} implementation of the \emph{GRDA}
classifier in the {\tt sparsediscrim} package, we set $\alpha_k = 1$, $k = 1,
\ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant
candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
cross-validation \citep{Hastie:2008dt}.

@@

For the microarray data sets, we computed error-rate estimates for the competing
classifiers after reducing the dimension of the data sets with feature selection
to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
\cite{Xu:2009fl}. To compare misclassification rates between different values of
$d$, we applied each value of $d$ considered to the same random allocation of
training and test data. Hence, for a given random partition, the set of features
selected for a small value of $d$ is a proper subset of the features selected
for a larger value of $d$. We report the average of the error-rate estimates for
each value of $d$ in Figure \ref{fig:error-rates}. Approximate standard errors
were no greater than \Sexpr{max(error_rates[["SE_Error"]])}.

The \emph{GRDA} classifier was superior in classification performance for the
majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, the
\emph{GRDA} classifier outperformed its competitors for all values of $d$
examined. For the Nakayama data set, the \emph{GRDA} classifier was outperformed
by the Dudoit, Pang, and Tong classifiers for small values of $d$ but attained
classification rates superior to its competitors for larger values of
$d$. Furthermore, Figure \ref{fig:error-rates} emphasizes two strengths of the
\emph{GRDA} classifier: reliability of the classifier for small $d$ and
stability of the classifier for large $d$. In contrast to the other classifiers,
the \emph{GRDA} classifier exhibited less variability its misclassification
estimates for small values of $d$. Also, the \emph{GRDA} classifier yielded
superior classification accuracy as $d$ increased and was hardly effected by the
inclusion of more variables in the data sets. We attribute the apparent
stability of the error-rate estimates of the \emph{GRDA} classifier for large
values of $d$ to its novel dimension-reduction technique.

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$. The Guo classifier
required a larger subset of the features to perform competitively in our study.

\[ \left[\text{Figure \ref{fig:error-rates} goes about here }\right]\]

Next, we examine the optimal estimates of the tuning parameters for the
\emph{GRDA} classifier. For brevity, we include only a graphical summary of the
optimal estimates of $\lambda$ for the Shipp data set in Figure
\ref{fig:optimal-shipp}. Similar figures for the other three data sets can be
found in the Supplementary Information. In Figure \ref{fig:optimal-shipp}, the
distribution of the optimal values of $\lambda$ selected through
cross-validation has a notable interesting bathtub shape, resembling a Beta
distribution with shape parameters $1/2$ and $1/2$. Specifically, the optimal
values of $\lambda$ are typically near 0 or near 1, corresponding to the
\emph{QDA} and \emph{LDA} classifiers, respectively. We noticed a similar
distribution of optimal values of $\lambda$ for the other three data sets.

For the Shipp data set with smaller values of $d$, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{GRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of the \emph{GRDA} classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious \emph{LDA} classifier as a special case.

\[ \left[\text{Figure \ref{fig:optimal-shipp} goes about here }\right]\]

@@

The analysis of high-dimensional microarray data sets is important to modern
biomedical studies and presents challenges with supervised classification
because the number of gene expressions greatly outnumbers the number of
collected responses. Consequently, several variants of the well-known \emph{LDA}
classifier have been proposed by employing for improved classification
performance. However, the \emph{LDA} classifier includes the strong assumption
that the covariance matrices of each class are equal. Ideally, we could model
the class covariance matrices as unrestricted and unequal in the \emph{QDA}
classifier, but it is well known to have a large number of parameters that must
be estimated and require a large number of samples, which are often unavailable
in the high-dimensional setting. With this mind, \cite{Friedman:1989tm} proposed
the \emph{RDA} classifier as a compromise between the \emph{LDA} and \emph{QDA}
classifiers, controlling the amount of pooling of the individual class
covariance-matrix estimators using the tuning parameter $\lambda$. However,
despite its success for data sets where $p > N$, the \emph{RDA} classifier
requires excessive calculations and a time-consuming model selection.

With this in mind, we have adapted the \emph{RDA} classifier for usage with
modern small-sample, high-dimensional data sets. Our model formulation
facilitates a rapid classification of high-dimensional data by employing a novel
dimension reduction technique. Moreover, our 10-fold cross-validation procedure
typically requires a small runtime for data sets having large $p$. When
preliminary variable selection is applied to the data as we performed in our
classification study, the time to perform model selection is nearly
instantaneous, confirming the computational complexity reported regarding the
operations employed in the \emph{GRDA} classifier. Additionally, we have
provided an implementation of the \emph{GRDA} classifier in the {\tt
  sparsediscrim} R package that is easy to use and available on CRAN.

We have parameterized the covariance-matrix regularization employed in the
\emph{GRDA} classifier to include a large family of shrinkage methods, such as
the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety
of other shrinkage methods studied by \cite{Ramey:2013ji} and
\cite{Xu:2009fl}. Our formulation of the \emph{GRDA} classifier in terms of the
nonzero eigenvalues of the pooled sample covariance matrix indicates that
eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the
similiarities in classification accuracy of several regularization methods
applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a
high-dimensional classification study. We have provided an invariance property
that extends results of \cite{Ji:2008wp} and \cite{Ye:2006jm}, who have provided
a similar invariance property for the \emph{LDA} decision rule. Additionally,
our model formulation improves the interpretation of the \emph{RDA} classifier
by explicitly defining the covariance-matrix estimator so that each of the $N$
observations contributes to the estimation of the individual class covariance
matrices.

In our simulation study, we compared the \emph{GRDA} classifier to several
recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
classifiers, each of which assumes independence of the features conditional on
class membership. Despite the rapid computational performance of diagonal
classifiers and the reduction in the number of parameters to estimate,
\cite{Mai:2012bf} have noted that diagonal classifiers can often yield inferior
classification performance to other classification methods. We have shown that
our proposed \emph{GRDA} classifier often yields superior classification
accuracy to the diagonal classifiers considered, confirming the assertion of
\cite{Mai:2012bf}. Furthermore, the \emph{GRDA} classifier is able to compete
with the diagonal classifiers in terms of speed by incorporating a novel
dimension-reduction technique. Although the diagonal classifiers are clearly
faster to compute, the differences in runtime for small-sample, high-dimensional
data sets is hardly noticeable.

@@

\begin{figure}
<<grda_contours, fig.width=8, fig.height=8, out.width='.49\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 1/4)
rda_contours(lambda = 3/4)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:grda-contours}
\end{figure}

\begin{figure}
<<error_rates>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(color = Classifier, linetype = Classifier), size = 1.5)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected") + ylab("Average Error Rate")
p
@ 
\caption{Average error-rate estimates of the competing classifiers as a function
  of $d$. Approximate standard errors are no greater than \Sexpr{max(error_rates[["SE_Error"]])}.}
\label{fig:error-rates}
\end{figure}
